---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# Global Air Quality Index Analysis

#### Program: Data Science

#### Affiliate: Binus University

#### Group Name:

1\. Andrew Jonathan - 2602072262

2\. Marvel Martawidjaja - 2602085656

3\. Melvern Amadio Hidayat - 2602072685\\

#### Description of Dataset:

This dataset provides comprehensive information on air quality indices across different cities in various countries. It encompasses data on key variables related to air pollution levels, as well as geographical coordinates for each city.

## **Exploratory Data Analysis**

```{r}
library(ggplot2)
library(dplyr)
library(dlookr)
library(Hmisc)
library(caret)
library(corrplot)
library(class)
library(pROC)
library(randomForest)
```

```{r}
df <- read.csv("C:/Users/T480/Downloads/archive/AQI and Lat Long of Countries.csv")
```

```{r}
head(df)
```

```{r}
str(df)
```

```{r}
summary(df)
```

```{r}
colSums(is.na(df))
```

```{r}
# Mengecek keberadaan duplikat pada seluruh baris dataset
duplikat <- duplicated(df)
# Mengembalikan nilai TRUE jika ada duplikat, FALSE jika tidak ada
any(duplikat)  
```

```{r}
# Plot AQI Value distribution
ggplot(df, aes(x = `AQI.Value`)) +
  geom_histogram(binwidth = 10, fill = "steelblue", color = "white") +
  labs(x = "AQI Value", y = "Count", title = "Distribution of AQI Values")
```

```{r}
# Plot AQI Category frequencies
df %>%
  count(`AQI.Category`) %>%
  ggplot(aes(x = reorder(`AQI.Category`, n), y = n, fill = `AQI.Category`)) +
  geom_bar(stat = "identity") +
  labs(x = "AQI Category", y = "Frequency", title = "AQI Category Frequencies") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
df %>%
  count(`AQI.Value`) %>%
  ggplot(aes(x = reorder(`AQI.Value`, n), y = n, color = `AQI.Value`)) +
  geom_point(size = 3) +
  labs(x = "AQI Category", y = "Frequency", title = "AQI Category Frequencies") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# Set the plot size
options(repr.plot.width = 16, repr.plot.height = 9)

# Create a color palette for AQI categories
aqi_colors <- c("Good" = "green", "Moderate" = "yellow", "Unhealthy for Sensitive Groups" = "orange",
                "Unhealthy" = "red", "Very Unhealthy" = "purple", "Hazardous" = "maroon")

# Create the scatterplot
ggplot(df, aes(x = lng, y = lat, color = `AQI.Category`)) +
  geom_point() +
  scale_color_manual(values = aqi_colors) +
  labs(x = "Longitude", y = "Latitude", title = "AQI Categories around the world", color = "AQI Category") +
  theme_bw()
```

```{r}
# Top 5 countries with highest AQI values
top_high_aqi <- aggregate(AQI.Value ~ Country, data = df, FUN = mean, na.rm = TRUE)
top_high_aqi <- top_high_aqi[order(top_high_aqi$AQI.Value, decreasing = TRUE), ]
top_high_aqi <- head(top_high_aqi, 5)

# Top 5 countries with lowest AQI values
top_low_aqi <- aggregate(AQI.Value ~ Country, data = df, FUN = mean, na.rm = TRUE)
top_low_aqi <- top_low_aqi[order(top_low_aqi$AQI.Value), ]
top_low_aqi <- head(top_low_aqi, 5)

# Display the top 5 countries with highest AQI values
cat("Top 5 countries with highest AQI values (worst air quality):\n")
print(top_high_aqi)

# Display the top 5 countries with lowest AQI values
cat("\nTop 5 countries with lowest AQI values (best air quality):\n")
print(top_low_aqi)
```

```{r}
# Create bar plot for top 5 countries with highest AQI values
ggplot(top_high_aqi, aes(x = reorder(Country, AQI.Value), y = AQI.Value)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(x = "Country", y = "Average AQI Value", title = "Top 5 Countries with Highest AQI Values") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()

# Create bar plot for top 5 countries with lowest AQI values
ggplot(top_low_aqi, aes(x = reorder(Country, AQI.Value), y = AQI.Value)) +
  geom_bar(stat = "identity", fill = "green") +
  labs(x = "Country", y = "Average AQI Value", title = "Top 5 Countries with Lowest AQI Values") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()
```

```{r}
# Encoding categorical variables
df$AQI.Category <- as.numeric(factor(df$AQI.Category))
df$CO.AQI.Category <- as.numeric(factor(df$CO.AQI.Category))
df$Ozone.AQI.Category <- as.numeric(factor(df$Ozone.AQI.Category))
df$PM2.5.AQI.Category <- as.numeric(factor(df$PM2.5.AQI.Category))

head(df)
```

```{r}
# Drop specified columns
df <- subset(df, select = -c(lat, lng, Country, City, NO2.AQI.Value, NO2.AQI.Category))

head(df)
```

```{r}
dim(df)
```

```{r}
# Select only numeric columns
numeric_cols <- sapply(df, is.numeric)
numeric_df <- df[, numeric_cols]

# Compute the correlation matrix
corr <- cor(numeric_df)

# Create heatmap
corrplot(corr, method = "color", type = "upper", tl.col = "black", tl.srt = 45)
```

```{r}
df$AQI.Category <- as.factor(df$AQI.Category)
```

## **Predictive Model**

```{r}
# Split the data into features (X) and target variable (y)
X <- df[, !(colnames(df) %in% c("AQI.Value"))]
y <- df$AQI.Value

# Split the data into training and testing sets
train_indices <- caret::createDataPartition(y, p = 0.8, list = FALSE)
X_train <- X[train_indices, ]
y_train <- y[train_indices]
X_test <- X[-train_indices, ]
y_test <- y[-train_indices]

# Check the dimensions of the training and testing sets
cat("Training set dimensions:", dim(X_train), "\n")
cat("Testing set dimensions:", dim(X_test), "\n")
```

```{r}
df

```

```{r}
# Train the model
model <- randomForest(y = y_train, x = X_train)

# Make predictions on the test set
predictions <- predict(model, newdata = X_test)

# Convert factor predictions to numeric
predictions <- as.numeric(as.character(predictions))

# Convert factor y_test to numeric
y_test <- as.numeric(as.character(y_test))

# Calculate MAE
mae <- caret::MAE(predictions, y_test)
cat("MAE:", mae, "\n")

# Calculate RMSE
rmse <- caret::RMSE(predictions, y_test)
cat("RMSE:", rmse, "\n")

# Calculate R-squared
rsquared <- caret::R2(predictions, y_test)
cat("R-squared:", rsquared, "\n")
```

```{r}
predictions
```

